On the Gravitational stability of gravito-turbulent accretion disks

### 2016-01-25 paper/application.tex
revising summary
version 2 in text, about to redo figs

\section{Application to protoplanetary disks}\label{2dppd}
We now apply our linear framework to assess the stability of realistic 
protoplanetary disks. We consider the gravito-turbulent disk models recently
developed by \cite[][hereafter \citetalias{rafikov15}]{rafikov15}. 
This 2D, Keplerian disk orbits a Solar 
mass star and is defined by the following parameters,

\begin{itemize}
  \item $\dot{M}$, the global radial mass accretion rate;
  \item $Q_0$, the value of the 2D Toomre parameter where the disk is
    gravito-turbulent;
  \item $T_\mathrm{irr}$, the irradiation or floor temperature;
  \item $\alpha_m$, the dimensionless viscosity associated with other
    sources of turbulence, such as magneto-rotational instabilities
    (MRI), see also \S\ref{visc_gi} and \S\ref{MHD}. 
\end{itemize} 
These are used in the statements of thermal equilibrium, mass and
angular momentum conservation, together with an opacity law, 
to construct a global disk model with surface density $\Sigma(R)$ and
temperature $T(R)$ where $R$ is the global cylindrical radius from the
star. \citepalias[See][for details.]{rafikov15} This gives the $Q(R)$
value required for input into our linear framework. We derive other
dimensionless parameters below.

@@

\subsection{PPD beta cooling}
Energy loss in \citetalias{rafikov15} is given by 
\begin{align}\label{real_cool}
  \Lambda = \frac{2\sigma}{f(\tau)}\left(T^4 - T_\mathrm{irr}^4\right)
\end{align}
per unit area, 
\begin{align}
  f(\tau) = \tau + \frac{1}{\tau}, \label{ftau} 
\end{align}
and
\begin{align}
  \tau = \kappa_d(T)\Sigma
\end{align}
is the optical depth and recall Eq. \ref{opacity_law} 
is our opacity model where $\kappa_d\propto T^b$ with $b=2$. 
Eq. \ref{ftau} accounts for cooling in the optically-thin ($\tau\ll
1$) and optically-thick ($\tau\gg1$) regimes.

Note that Eq. \ref{real_cool} is a beta cooling prescription because
it is an explicit function of the thermodynamic states. However, we
formulated the linear problem with beta cooling of the functional form
given by Eq. \ref{beta_cool}. In order to adapt the developed
framework to the above PPD beta cooling function, we
need to identify the $\beta$ and $\theta$ parameters that enter the 
linearized equations (Eq. \ref{linear_beta}).

@@

Eq. \ref{real_beta} still represents a physical (dimensionless) cooling
time, and is consistent with previous definitions of realistic cooling
times within factors of order unity \citep[e.g.][their
Eq. 2]{kratter10}.

However, $\theta$ is now an effective 
irradiation level. 
%no longer a direct measure of
%irradiation. 
Eq. \ref{real_theta} shows that $\theta$ is related to the true
irradation temperature $\tirr$ through the function $g$ given by 
Eq. \ref{g_def}. This is
important because a physically non-irradiated disk with $\tirr=0$ will
still have a non-zero effective irradiation, $\theta\neq0$. 
%This results from the fact that PPD cooling generally depends on both
%$P$ and $\Sigma$, but in the absence of irradiation standard beta
%cooling only depends on $P$.

@@

%the gammie-type cooling function does not represent realistic
%cooling. because real cooling depends on sigma and pressure but gammie
%cooling only depends on pressure

@@

PPDs become irradiation-dominated at large distances from the star,
where $T\to\tirr$, or $\theta\to 1$. {\bf ref?}
Then the first inequality in Eq. \ref{ppd_invisc_cond} relaxes to
$\gamma > 1$ in the outer disk, which is always satisfied.  
On the other hand, numerical simulations of gravito-turbulence
show that $1\lesssim Q \lesssim 2$ \citep{gammie01,rice11}, and 
the second inequality is generally satisfied. Taken together, this
means that in the outer regions of a realistic PPD, cooling itself
cannot cause instability.

\subsection{Example 2D calculation}\label{pp2d_example}
We consider a model with $\dot{M} = 10^{-6}M_\sun\,\yr^{-1}$, $Q_0=1.5$, 
$\tirr=10\mathrm{K}$, and $\alpha_m=10^{-3}$. 
We adopt the opacity scale $\kappa_{d0} =
5\times10^{-4}\mathrm{cm}^2\,\mathrm{g}^{-1}\,\mathrm{K}^{-2}$  as in
\citetalias{rafikov15}. We use $\gamma=1.6$, approximately applicable
to molecular gas at low temperatures and used by other authors
\citep{rice11,baehr15}. Then the global inviscid stability condition
(Eq. \ref{ppd_invisc_cond}) is satisfied. 
%for outer disk choice of gamma doesn't matter 
%avoid complications from over-stable thermal modes 
Fig. \ref{rafikov_model} shows the equilirbium 
disk profile in terms of $Q$, $\alpha$, $\beta$, and $\theta$.  
These are set as input to the 2D dispersion relation
(Eq. \ref{thindisk}).

Fig. \ref{rafikov_growth} show growth timescales and
optimum wavenumbers for viscous GI in this disk model.  
We also plot analytic estimates based on Eq. \ref{gammie_smallk} (instead of
Eq. \ref{gammie_maxrate_simple} since here $\theta\sim 1$) 
which gives the optimum wavenumber as
\begin{align}
  |K| = \frac{3}{2\theta Q},
\end{align}
with growth rate
\begin{align} 
  S = \frac{27\alpha}{16\theta^3Q^4}. 
\end{align}
These are similar to the isothermal results of
\citet[][their Eq. 19 and 21, respectively]{sterzik95}, and identical if one
takes $\theta=1$. %showing that irradiation makes the disk behave like
                  %isothermal disk

@@

%{\bf despite apply beta cooling results to realistic cooling disks}
%Our result suggest an alternative interpretation of disk fragmentation
%is where the gravito-turbulent state becomes dynamically unstable due to
%both viscosity and self-gravity. 
%{\bf note: disk might be stablized by extremely large alpha at large
%  distances but `turbulent' framework no longer valid }

@@

Fig. \ref{rafikov_growth3d} show the growth rates and most unstable
wavenumber for $R\in[10,100]$AU; along with corresponding 2D 
results with softened self-gravity. There is a close match for
$R\lesssim60$AU where the disk is optically thick ($\tau\gtrsim
1$) and both models apply. However, beyond $60$AU where the disk
becomes optically-thin, radiative diffusion (the 3D curve) is not
valid and under-estimates the growth rates. %; while the $f(\tau)$
%function approximately captures opticall-thin cooling in the 2D case.  
Nevertheless, the comparison shows that the transition radius
$R_t\sim60$AU can be calculated within the 2D framework.  
%This is not surprising since the
%instability is fundamentally driven by viscous and self-gravity forces in the
%momentum equations, not the energy equation.

>>>>

\section{Application to protoplanetary disks}\label{2dppd}
We now apply our linear framework to assess the stability of  
protoplanetary disks. We consider the gravito-turbulent disk models recently
developed by \cite[][hereafter \citetalias{rafikov15}]{rafikov15}. 
This 2D, Keplerian disk orbits a Solar 
mass star and is defined by the following parameters,

\begin{itemize}
  \item $\dot{M}$, the global radial mass accretion rate;
  \item $Q_0$, the value of the 2D Toomre parameter where the disk is
    gravito-turbulent;
  \item $T_\mathrm{irr}$, the irradiation or floor temperature;
  \item $\alpha_m$, the dimensionless viscosity associated with other
    sources of turbulence, such as magneto-rotational instabilities
    (MRI, see also \S\ref{visc_gi} and \S\ref{MHD}). 
\end{itemize} 
These are used in the statements of thermal equilibrium, mass and
angular momentum conservation, together with an opacity law, 
to construct a global disk model with surface density $\Sigma(R)$ and
temperature $T(R)$ where $R$ is the global cylindrical radius from the
star. \citepalias[See][for details.]{rafikov15} This gives the $Q(R)$
value required for input into our linear framework. We derive other
dimensionless parameters below.

@@

\subsection{PPD beta cooling}\label{ppd_cooling}
Energy loss in \citetalias{rafikov15} is given by 
\begin{align}\label{real_cool}
  \Lambda = \frac{2\sigma}{f(\tau)}\left(T^4 - T_\mathrm{irr}^4\right)
\end{align}
per unit area, 
\begin{align}
  f(\tau) = \tau + \frac{1}{\tau}, \label{ftau} 
\end{align}
and
\begin{align}
  \tau = \kappa_d(T)\Sigma
\end{align}
is the optical depth. Recall Eq. \ref{opacity_law} 
is our opacity model where $\kappa_d\propto T^b$ with $b=2$. 
Eq. \ref{ftau} accounts for cooling in the optically-thin ($\tau\ll
1$) and optically-thick ($\tau\gg1$) regimes.

Note that Eq. \ref{real_cool} is a beta cooling prescription because
it is an explicit function of the thermodynamic states. However, we
formulated the linear problem with the standard beta cooling function 
given by Eq. \ref{beta_cool}. In order to adapt the existing  
framework to the above PPD beta cooling function, we 
need to identify the $\beta$ and $\theta$ parameters that enter the 
linearized equations (Eq. \ref{linear_beta}).

@@

Eq. \ref{real_beta} still represents a physical cooling 
time, and is consistent with previous definitions within factors of
order unity \citep[e.g.][their Eq. 2]{kratter10}.

However, $\theta$ is now an effective 
irradiation level. 
Eq. \ref{real_theta} shows that $\theta$ is related to the true
irradation temperature $\tirr$ through the function $g$ given by 
Eq. \ref{g_def}. This is
important because a physically non-irradiated disk with $\tirr=0$ will
still have a non-zero effective irradiation, $\theta\neq0$. 
%This results from the fact that PPD cooling generally depends on both
%$P$ and $\Sigma$, but in the absence of irradiation standard beta
%cooling only depends on $P$.

@@

@@

PPDs become irradiation-dominated at large distances from the star,
where $T\to\tirr$ and $\theta\to 1$. {\bf ref?}
Then the first inequality in Eq. \ref{ppd_invisc_cond} relaxes to
$\gamma > 1$ in the outer disk, which is always satisfied.  
On the other hand, numerical simulations of gravito-turbulence
show that $1\lesssim Q \lesssim 2$ \citep{gammie01,rice11}, and 
the second inequality is generally satisfied. Taken together, this
means that in the outer regions of a realistic PPD, cooling itself
cannot cause instability.

\subsection{Example 2D calculation}\label{pp2d_example}
We consider a disk  model with $\dot{M} = 10^{-6}M_\sun\,\yr^{-1}$, $Q_0=1.5$, 
$\tirr=10\mathrm{K}$, and $\alpha_m=10^{-3}$. 
We adopt the opacity scale $\kappa_{d0} =
5\times10^{-4}\mathrm{cm}^2\,\mathrm{g}^{-1}\,\mathrm{K}^{-2}$  as in
\citetalias{rafikov15}. We use $\gamma=1.6$, approximately applicable
to molecular gas at low temperatures and used by other authors
\citep{rice11,baehr15}. Then the global inviscid stability condition
(Eq. \ref{ppd_invisc_cond}) is satisfied. %which simplifies the problem
%for outer disk choice of gamma doesn't matter 
%avoid complications from over-stable thermal modes 
Fig. \ref{rafikov_model} shows the equilirbium 
disk profile in terms of $Q$, $\alpha$, $\beta$, and $\theta$.  
These are set as input to the 2D dispersion relation
(Eq. \ref{thindisk}).

Fig. \ref{rafikov_growth} show growth timescales and
optimum wavenumbers for viscous GI in this disk model.  
We also plot analytic estimates based on Eq. \ref{gammie_smallk} (instead of
Eq. \ref{gammie_maxrate_simple} since here $\theta\sim 1$) 
which gives the optimum wavenumber and growth rates as
\begin{align}
  |K| = \frac{3}{2\theta Q}, \quad
%\end{align}
%with growth rate
%\begin{align} 
  S = \frac{27\alpha}{16\theta^3Q^4}. 
\end{align}
These are similar to the isothermal results of
\citet[][their Eq. 19 and 21, respectively]{sterzik95}, and identical if one
takes $\theta=1$. %showing that irradiation makes the disk behave like
                  %isothermal disk

@@

@@

Fig. \ref{rafikov_growth3d} show the growth rates and most unstable
wavenumber for $R\in[10,100]$AU; along with corresponding 2D 
results with softened self-gravity. There is a close match for
$R\lesssim60$AU where the disk is optically thick ($\tau\gtrsim
1$) and both models apply. However, beyond $60$AU where the disk
becomes optically-thin, radiative diffusion (the 3D curve) is not
valid and under-estimates the growth rates. %; while the $f(\tau)$
%function approximately captures opticall-thin cooling in the 2D case.  
However, the comparison shows that the transition radius
$R_t\sim60$AU can be correctly calculated within the 2D framework.  
%notice the small softenings used 
%This is not surprising since the
%instability is fundamentally driven by viscous and self-gravity forces in the
%momentum equations, not the energy equation.

### 2016-01-25 paper/linear3d.tex
regular update
revising sec 4, minor edits to prev secs
ready to revise applications
version 2 in text, about to redo figs

%On the other hand, if a time-independent heat source is invoked, then
%we may choose $\alpha, \beta$ and $\theta$ separately.  
\section{Linear problem}\label{linear}
We consider infinitestimal axisymmetric Eulerian perturbations in the
form \begin{align}
  \delta \rho= \widetilde{\delta \rho}(z)\exp{\left(\ii kx +
    s t\right)},  
\end{align}
and similarly for other variables. Here, $k$ 
is a real horizontal wavenumber and $s$ is a (generally) complex
growth rate. 
%The gravitational potential perturbation $\delta \Phi$ depends on $z$
%but only its midplane value $\delta \Phi(z=0)$ is
%used in the dynamical equations, which has the same form as above. 
For clarity, hereafter we drop the tilde.  
%and note that equilibrium
%quantities without the $\delta$ prefix.

The linearized continuity, momentum and energy equations are
\begin{align}
   &s \delta \rho = - \ii k \rho\delta v_x -
  \left(\rho\delta v_z\right)^\prime \label{lin_mass} \\ 
   &s  \delta v_x = - \ii k\frac{\delta P}{\rho} - \ii k \delta 
  \Phi + 2\Omega \delta v_y + \delta F_x,\\
   &s \delta v_y = (q-2)\Omega\delta v_x + \delta F_y,\\
%   &s \delta v_z = - \left(\frac{\delta P}{\rho}\right)^\prime -
%  \delta\Phi^\prime +
%  \left(\ln{\rho}\right)^\prime\left(c_s^2\frac{\delta\rho}{\rho} -
%  \frac{\delta P}{\rho}\right) + \delta F_z,\\
  &s \delta v_z =
  \left(\ln{P}\right)^\prime c_s^2\frac{\delta\rho}{\rho} -
  \left(\frac{\delta P}{\rho}\right)^\prime -
  \left(\ln{\rho}\right)^\prime\frac{\delta P}{\rho} + \delta F_z, \\ 
  &s \frac{\delta P}{\rho} = 
%- \ii \gamma k \frac{c_s^2}{\Gamma}\delta v_x 
 - \ii k \gamma c_s^2\delta v_x
%- c_s^2\left[
%  \left(\ln{\rho}\right)^\prime\delta v_z + 
%  \frac{\gamma}{\Gamma}\delta v_z^\prime\right] \notag\\
- c_s^2\left[
 \left(\ln{P}\right)^\prime\delta v_z + 
  \gamma\delta v_z^\prime\right] \notag\\
   & \phantom{s \frac{\delta P}{\rho}=}
  + (\gamma-1)\frac{\delta\mathcal{H}_\mathrm{visc}}{\rho} 
  -(\gamma-1)\frac{\delta\Lambda}{\rho}
%  - \frac{1}{t_c}\left(\frac{\delta P}{\rho} -
%  \frac{\theta}{\Gamma}c_s^2\frac{\delta\rho}{\rho}\right) 
  \label{lin_energy},\\
  &\delta\Phi^{\prime\prime} - k^2\delta\Phi =
  \frac{\Omega^2}{Q_\mathrm{3D}}\left(\frac{\delta\rho}{\rho_0}\right)\label{lin_gravity},  
\end{align}
with
\begin{align}
  \delta F_x =& \nu \left[\delta v_x^{\prime\prime} + 
    \left(\ln{\rho\nu}\right)^\prime \delta v_x^\prime - 
     \frac{4}{3}k^2\delta v_x \right] - \nu_bk^2\delta v_x \notag\\
   &+ \ii \nu k \left[\frac{1}{3}\delta v_z^\prime+ 
    \left(\ln{\rho\nu}\right)^\prime\delta v_z\right] + \ii k \nu_b
  \delta v_z^\prime,\\
  \delta F_y = & \nu  \left[\delta v_y^{\prime\prime} + 
    \left(\ln{\rho\nu}\right)^\prime \delta v_y^\prime - 
     k^2\delta v_y \right] - \ii\nu kq\Omega\delta\ln{\rho\nu},\\
  \delta F_z =& \nu\left[\frac{4}{3}\delta v_z^{\prime\prime} + 
    \frac{4}{3}\left(\ln{\rho\nu}\right)^\prime \delta v_z^\prime - 
    k^2\delta v_z \right] \notag\\
  & +\nu_b\left[\delta v_z^{\prime\prime} + \left(\ln{\rho\nu_b}\right)^\prime\delta v_z^\prime\right]\notag\\
  & + \ii \nu k  \left[\frac{1}{3}\delta v_x^\prime -  
    \frac{2}{3}\left(\ln{\rho\nu}\right)^\prime\delta v_x\right]\notag\\
  &+\ii \nu_b k \left[\delta v_x^\prime + \left(\ln{\rho\nu_b}\right)^\prime\delta v_x\right],\\
%  &\delta f_x = -\frac{4}{3} \nu k^2\Sigma \delta v_x,\\
%  &\delta f_y = -\ii q k \Omega \delta(\nu\Sigma) - \nu k^2
%  \Sigma\delta v_y,\\
%  &\delta Q_\mathrm{visc} = (q\Omega)^2\delta(\nu\Sigma) - 2 \ii \nu 
%  k q\Omega\Sigma\delta v_y,\\
%  &\delta(\nu\Sigma) =
%  \frac{\alpha}{\Omega}\left[c_s^2(1+\mu)\delta\Sigma + \lambda\delta P\right].
%  &\delta f_y = - \ii \nu q k \Omega \delta \Sigma - \nu k^2\Sigma\delta
%  v_y,\\
%  &\delta Q_\mathrm{visc} = \nu
%  \left(q\Omega\right)^2\delta\Sigma - 2\ii\nu q k \Omega\Sigma\delta
%  v_y. 
  \frac{\delta\mathcal{H}_\mathrm{visc}}{\rho}=& \nu (q\Omega)^2
  \delta\ln{\rho\nu} - 2\ii\nu k q \Omega \delta v_y,\\
  \delta\ln{\rho\nu} =& (1+\mu)\frac{\delta\rho}{\rho} +
  \frac{\lambda}{c_s^2}\frac{\delta P}{\rho}, \label{linear_beta_cool}
\end{align}
where $^\prime$ denotes $d/dz$, and the 3D self-gravity parameter is
\begin{align}
  Q_\mathrm{3D} \equiv \frac{\Omega^2}{4\pi G \rho_0}. 
\end{align}
The linearized cooling functions $\delta\Lambda$ are given below. 
Eq. \ref{lin_mass}---\ref{lin_gravity}, supplemented with appropriate
boundary conditions, constitutes an eigenvalue problem for the growth
rate $s$.

>>>>

\section{Linear problem}\label{linear}
We consider infinitestimal axisymmetric Eulerian perturbations in the
form \begin{align}
  \delta \rho= \widetilde{\delta \rho}(z)\exp{\left(\ii kx +
    s t\right)},  
\end{align}
and similarly for other variables. Here, $k$  is an input real horizontal
wavenumber and $s$ is a (generally) complex growth rate.  
For simplicity, hereafter we drop the tilde.

The linearized continuity, momentum and energy equations are
\begin{align}
   &s \delta \rho = - \ii k \rho\delta v_x -
  \left(\rho\delta v_z\right)^\prime \label{lin_mass} \\ 
   &s  \delta v_x = - \ii k\frac{\delta P}{\rho} - \ii k \delta 
  \Phi + 2\Omega \delta v_y + \delta F_x,\\
   &s \delta v_y = (q-2)\Omega\delta v_x + \delta F_y,\\
%   &s \delta v_z = - \left(\frac{\delta P}{\rho}\right)^\prime -
%  \delta\Phi^\prime +
%  \left(\ln{\rho}\right)^\prime\left(c_s^2\frac{\delta\rho}{\rho} -
%  \frac{\delta P}{\rho}\right) + \delta F_z,\\
  &s \delta v_z =
  \left(\ln{P}\right)^\prime c_s^2\frac{\delta\rho}{\rho} -
  \left(\frac{\delta P}{\rho}\right)^\prime -
  \left(\ln{\rho}\right)^\prime\frac{\delta P}{\rho} + \delta F_z, \\ 
  &s \frac{\delta P}{\rho} = 
%- \ii \gamma k \frac{c_s^2}{\Gamma}\delta v_x 
 - \ii k \gamma c_s^2\delta v_x
%- c_s^2\left[
%  \left(\ln{\rho}\right)^\prime\delta v_z + 
%  \frac{\gamma}{\Gamma}\delta v_z^\prime\right] \notag\\
- c_s^2\left[
 \left(\ln{P}\right)^\prime\delta v_z + 
  \gamma\delta v_z^\prime\right] \notag\\
   & \phantom{s \frac{\delta P}{\rho}=}
  + (\gamma-1)\frac{\delta\mathcal{H}_\mathrm{visc}}{\rho} 
  -(\gamma-1)\frac{\delta\Lambda}{\rho}
%  - \frac{1}{t_c}\left(\frac{\delta P}{\rho} -
%  \frac{\theta}{\Gamma}c_s^2\frac{\delta\rho}{\rho}\right) 
  \label{lin_energy},\\
  &\delta\Phi^{\prime\prime} - k^2\delta\Phi =
  \frac{\Omega^2}{Q_\mathrm{3D}}\left(\frac{\delta\rho}{\rho_0}\right)\label{lin_gravity},  
\end{align}
where $^\prime$ denotes $d/dz$. The perturbed viscous forces are 
\begin{align}
  \delta F_x =& \nu \left[\delta v_x^{\prime\prime} + 
    \left(\ln{\rho\nu}\right)^\prime \delta v_x^\prime - 
     \frac{4}{3}k^2\delta v_x \right] - \nu_bk^2\delta v_x \notag\\
   &+ \ii \nu k \left[\frac{1}{3}\delta v_z^\prime+ 
    \left(\ln{\rho\nu}\right)^\prime\delta v_z\right] + \ii k \nu_b
  \delta v_z^\prime,\\
  \delta F_y = & \nu  \left[\delta v_y^{\prime\prime} + 
    \left(\ln{\rho\nu}\right)^\prime \delta v_y^\prime - 
     k^2\delta v_y \right] - \ii\nu kq\Omega\delta\ln{\rho\nu},\\
  \delta F_z =& \nu\left[\frac{4}{3}\delta v_z^{\prime\prime} + 
    \frac{4}{3}\left(\ln{\rho\nu}\right)^\prime \delta v_z^\prime - 
    k^2\delta v_z \right] \notag\\
  & +\nu_b\left[\delta v_z^{\prime\prime} + \left(\ln{\rho\nu_b}\right)^\prime\delta v_z^\prime\right]\notag\\
  & + \ii \nu k  \left[\frac{1}{3}\delta v_x^\prime -  
    \frac{2}{3}\left(\ln{\rho\nu}\right)^\prime\delta v_x\right]\notag\\
  &+\ii \nu_b k \left[\delta v_x^\prime + \left(\ln{\rho\nu_b}\right)^\prime\delta v_x\right],
%  &\delta f_x = -\frac{4}{3} \nu k^2\Sigma \delta v_x,\\
%  &\delta f_y = -\ii q k \Omega \delta(\nu\Sigma) - \nu k^2
%  \Sigma\delta v_y,\\
%  &\delta Q_\mathrm{visc} = (q\Omega)^2\delta(\nu\Sigma) - 2 \ii \nu 
%  k q\Omega\Sigma\delta v_y,\\
%  &\delta(\nu\Sigma) =
%  \frac{\alpha}{\Omega}\left[c_s^2(1+\mu)\delta\Sigma + \lambda\delta P\right].
%  &\delta f_y = - \ii \nu q k \Omega \delta \Sigma - \nu k^2\Sigma\delta
%  v_y,\\
%  &\delta Q_\mathrm{visc} = \nu
%  \left(q\Omega\right)^2\delta\Sigma - 2\ii\nu q k \Omega\Sigma\delta
%  v_y. 
\end{align}
and the perturbed viscous heating is given by 
\begin{align}
  \frac{\delta\mathcal{H}_\mathrm{visc}}{\rho}=& \nu (q\Omega)^2
  \delta\ln{\rho\nu} - 2\ii\nu k q \Omega \delta v_y,\\
  \delta\ln{\rho\nu} =& (1+\mu)\frac{\delta\rho}{\rho} +
  \frac{\lambda}{c_s^2}\frac{\delta P}{\rho}. \label{linear_beta_cool}
\end{align}
In Eq. \ref{lin_gravity}, the 3D self-gravity parameter is
\begin{align}\label{Q3d_def}
  Q_\mathrm{3D} \equiv \frac{\Omega^2}{4\pi G \rho_0}. 
\end{align}
The linearized cooling functions $\delta\Lambda$ are given below. 
Eq. \ref{lin_mass}---\ref{lin_gravity}, supplemented with appropriate
boundary conditions, constitutes an eigenvalue problem for the growth
rate $s$.
